---
title: "covid 19"
author: "Alex Moreno"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$:There is no correlation between the two. The sex of a person does not determine whether they have a higher/lower likelihood of dying or not. You have the same likelihood of death based off of gender. Dying is independent. 

$H_A$: There is a correlation between the two. The sex of a person does determine whether they ave a higher/lower likelihood of dying or not. You don't have the same likelihood of death based off of gender. Dying is dependent. 



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

# Methods

Cat~Cat
Bar plot will be used as well as chi-squared test and fisher exact test

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")
```

Frequency chart shows that mostly bboth stayed alive. Very little deaths in frequency. More females alive than males. The percent chart shows the death in Men to be more than females. This is accurate because in the frequency chart Males did have more deaths than females. Even if the numbers are so low. The percentage rate shows that Males tend to die more than woman from the virus.

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The data shows that 36 women  out of 100 people died out in the dead group. While 64 men out of 100 people died out in the dead group. Men seem to be dying more than females.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Null hypothesis is rejected in favor of the alternative. The gender does play a role. Males tend to die more than females. P-value is less than 0.05, it is 0.001 for the chi-squared test. P-value for fisher exact test is 0.0009 which is also less than 0.05. The alternative hypothesis is in favor. Odds ratie is at 2.3. Males are more likely to die from the virus.


## Conclusions

Is death dependent or independent of gender? From the data and analysis we can conclude that it is dependent. From the values given, Males tend to die more than Females. Both chi-0squared and fisher exact test give p-values of less than 0.05. The null hypothesis is rejected in favor of the alternative. Males will die from covid more than females will. 

